\documentclass[10pt,twocolumn]{article}
\usepackage{amsmath,amssymb,fullpage,fancyhdr}
\setlength{\parindent}{0in}

\begin{document}
\begin{center}\textsc{greedy algorithms}\end{center} 
{\bf kruskal:}  repeatedly add the next lightest edge to $G'$ that does not create a cycle. makeset(x), find(x): identify x's connected component (labeled by root), union(x,y): merge connected components. For all $x\in V: makeset(x)$. X=$\{\}$ (empty set). for all $e=(x,y)$ (in increasing order of weight), if $find(x)\neq find(y): X=X \cup \{e\}, union(x,y)$. Cost = sorting edges + $2\cdot |E|\cdot$cost(find) + $(|V|-1)\cdot$cost(union) + $|V|\cdot$cost(makeset).

{\bf prim}: similar to Dijkstra. Start form any vertex, find the lightest edge extending from your current tree and added that tree and ending vertex.

{\bf horn formulas:} $(x_1\land x_2 \land \cdots \land x_k)\Rightarrow x_{k+1}$ OR $(\bar{x_1}\lor \bar{x_2}\lor \cdots \lor \bar{x_k}\lor x_{k+1})$. Start with all vars = false. While there is an unsatisfied implication, set the implied (rightmost) variable to true. Return the truth assignments
	
	
\begin{center}\textsc{dynamic programming algs}\end{center} 
{\bf longest increasing subsequence: $O(n^2)$}:
$L(j)$=len of longest subsequence ending @ j. $L(j)=1+\max\{L(i)\}$ where $i<j,a_i < a_j \rightarrow (i,j)$ is an edge. Return $\max\{L(i) for i\leq n\}$, and loop from $i=n\rightarrow 0$

{\bf edit distance: $O(m\cdot n)$}: $E[m,n]=$min num of edits to change $x[1\cdots m]$ to $y[1\cdots n]$. E[m,n] = \[\min\left\{ \begin{array}{l l} \cdots x[m] , \cdots - \\ \mbox{delete x[m]: $E[m,n]=1+E[m-1,n]$} \\ \cdots -, \cdots y[n] \\ \quad \mbox{insert y[n]: $E[m,n]=1+E[m,n-1]$} \\ \cdots x[m], \cdots y[n] \\ \quad \mbox{change x[m] to y[n]:}\\ \mbox{ $E[m,n]=E[m-1,n-1] +$ (1 if diff, 0 else)}  \end{array} \right. \]

Initialize $i=0\cdots m: E[i,0] = i \quad j=0\cdots n: E[0,j]=j$. Loop $i=1\cdots m, j=1\cdots n, E[i,j]$

{\bf Knapsack: $O(W_{\max}\cdot n)$}: $K(w)=$max value of weight $w\leq W_{\max}$. $K(w)=\displaystyle\max_{w_i\leq w}\{v_i+K(w-w_i)\}$ $K(0)=0$. Loop $w=1\cdots W_{\max}$

{\bf Chain Matrix Multiply: $O(n^3)$} : $C(i,j)$ = cost of best solution to multiplying $A_i\cdots A_j$. \[C(i,j)=\displaystyle\min_{i\leq k\leq j}\{ C(i,k)+C(k+1,j) + m_{i-1}\cdot m_k \cdot m_j\}\] Solve in order of increasing subproblem length: for $i=1\cdots n-1: C(i,i=0)$. for $s=1\cdots n-1$, for $i=1\cdots n-s: j=i+s; $ update. Return $C(1,n)$.

{\bf Shortest path (all pairs of vertices): $O(|V|\cdot(|V|+|E|))$}: $dist(v,i)=$ dist from $s$ to $v$ using $i$ edges. $dist(v,i)=\displaystyle\min_{e=(w,v)}\{len(e)+dist(w,i-1)\}$. For all $v\in V \quad dist(v,0)=\infty, dist(s,0)=0$. for $i=1\cdot |V|$, for all $v\in V$, update.

{\bf Independent Sets: $O(|V|+|E|)$}: $I(u)=$size of largest independent set in subtree rooted at $u$. \[I(u) = \max\left\{ \begin{array}{l l} 1+\displaystyle\sum_{grandchildren} I(g_i) & \quad \mbox{if $u\in I(n)$} \\ \displaystyle\sum_{children} I(c_i) & \quad \mbox{u isn't} \\ \end{array} \right. \]

DFS traversal: postvisit[u] = update (does $u$ after children of $u$ are done)

{\bf Travelling Salesman: $O(2^nn^2)$} 
Subset $S \leq V$, $\{1\} \in S$.

for $j\in S, C(S,j)=$length of shortest path that starts at 1, ends at $j$, and visits each vertex in S once. If $|S| > 1, C(S,1)=\infty$.

update: $C(S,j)=\displaystyle\min_{i\in S, i\neq j, (i,j)\in E}\{C(S-\{j\},i)+len(i,j)\}$

For $S=2\cdots n: \mbox{for all sizes of S,}$ for all subsets $S\leq \{1\cdots n\}$ of size $s$ (including $\{i\}$): $C(S,1)=\infty$ for all $j\in S, j\neq 1, update$

\end{document}